Making Sequential Programs Deterministic

### 2019-04-04 java-design.tex
structuring section 3
Finished with partial pass through section 2
benchmark example

\Cref{code-determinism} presents some of the JDK methods that we have annotated with our determinism types and give examples of
client code that would produce errors at compile time.
\begin{figure}
    \begin{verbatim}
    // Annotated JDK methods
    public class Random implements java.io.Serializable {
        public @NonDet Random() {}
    }
    public class PrintStream extends FilterOutputStream 
        implements Appendable, Closeable {
        public void println(@Det Object x) {}
    }

    // Client code
    class Client {
        void test() {
            @Det double d = Math.random(); // Error - subtyping rules violated.
            @NonDet double nd = Math.random(); // No error.
            System.out.println(nd); // Error - println takes @Det arguments.
        }
    }
    \end{verbatim}
    \caption{Example: Errors detected by \theDeterminismChecker.}
\todo{Use real examples of bugs from case studies, throughout.  Artificial
  examples are less compelling (and lead the reader to wonder whether no
  real example exist).  Put the first example of an error in a figure in
  Section 1, to help with motivating the problem.}
    \label{code-determinism}
\end{figure}

@@

@@

\subsection{Maps}\label{sec:maps-java}

We prohibit the creation of a \codeid{@Det HashMap} or an \codeid{@OrderNonDet TreeMap} in our determinism checker.
\Cref{fig-creation-rules} formalizes these rules. Similar constraints apply to \<HashSet>s and \<TreeSet>s which are
also expressed by the last two rules of \cref{fig-creation-rules}.

>>>>

\Cref{code-determinism} presents some of the JDK methods that we have annotated with our determinism types and give examples of
client code that would produce errors at compile time.
%\begin{figure}
%    \begin{verbatim}
%    // Annotated JDK methods
%    public class Random implements java.io.Serializable {
%        public @NonDet Random() {}
%    }
%    public class PrintStream extends FilterOutputStream 
%        implements Appendable, Closeable {
%        public void println(@Det Object x) {}
%    }
%    
%    // Client code
%    class Client {
%        void test() {
%            @Det double d = Math.random(); // Error - subtyping rules violated.
%            @NonDet double nd = Math.random(); // No error.
%            System.out.println(nd); // Error - println takes @Det arguments.
%        }
%    }
%    \end{verbatim}
%    \caption{Example: Errors detected by \theDeterminismChecker.}
%\todo{Use real examples of bugs from case studies, throughout.  Artificial
%  examples are less compelling (and lead the reader to wonder whether no
%  real example exist).  Put the first example of an error in a figure in
%  Section 1, to help with motivating the problem.}
%    \label{code-determinism}
%\end{figure}

\begin{figure}
    \begin{verbatim}
    // Annotated JDK methods
    public interface Map<K,V> {
        @PolyDet Set<Map.Entry<K, V>> entrySet(@PolyDet Map<K,V> this);
    }
    public interface Iterator<E> {
        @PolyDet("up") E next(@PolyDet Iterator<E> this);
    }

    // Client code
    public class MapUtils {
        public static <K extends @PolyDet Object, V extends @PolyDet Object> 
                                @Det String toString(@PolyDet Map<K,V> map) {
            ...
            for(Map. @Det Entry<K,V> entry : map.entrySet()) { ... }
        }
    }
    [ERROR] MapUtils.java:[20,50] [enhancedfor.type.incompatible] 
    incompatible types in enhanced for loop.
    found : @PolyDet Entry<K extends @PolyDet Object,V extends @PolyDet Object>
    required: @Det Entry<K extends @PolyDet Object,V extends @PolyDet Object>
    \end{verbatim}
    \caption{Example: Error detected by \theDeterminismChecker in \<scribe-java>~\cite{nondex}.}
%    \todo{Use real examples of bugs from case studies, throughout.  Artificial
%        examples are less compelling (and lead the reader to wonder whether no
%        real example exist).  Put the first example of an error in a figure in
%        Section 1, to help with motivating the problem.}
    \label{code-determinism}
\end{figure}

@@

\subsection{Java collection types}\label{sec:maps-java}

We prohibit the creation of a \codeid{@Det HashMap} or an \codeid{@OrderNonDet TreeMap} in our determinism checker.
\Cref{fig-creation-rules} formalizes the type wellformedness rules for maps and sets. Similar constraints apply to \<HashSet>s and \<TreeSet>s which are
also expressed by the last two rules of \cref{fig-creation-rules}.

@@
